---
title: "covid 19"
author: "Dom Ciriello"
date: "3/18/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$ COVID-19 deaths in Korea, Jan-May 2020 are unrelated to sex.
$H_A$ There is a relationship between Korean COVID deaths and sex of the patient from Jan-May 2020.

Table of patients:
```{r}

kRdata$Sex = as.factor(kRdata$Sex)
kRdata$result = as.factor(kRdata$result)

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```
The table above shows that 19 more males died during that time period than women. More research is required to make a determination on the null hypothesis.

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Bar Chart 1 clearly shows the difference in total number of living women vs men. Bar chart 2 shows a clear difference in death percentage. Men showed 65% while women only showed 35%.

## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This table shows that Korean men were 2.14 times more likely than women to die from COVID-19 from Jan-May 2020 if you divide 3 by 1.4.

## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

## Conclusions

Since the p-value is .001, we reject the null hypothesis. There is a very small likelihood that the results happened purely by chance. There is an actual difference between male and female deaths from COVID-19 in Korea from Jan-May 2020 (65% vs 35%) and men were 2.3 times more likely to die than women.